
\documentclass[12pt]{article}
 
\usepackage[margin=1in]{geometry} 
\usepackage{amsmath,amsthm,amssymb}
\usepackage{pifont}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{array}
 
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
 
\newenvironment{theorem}[2][Theorem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{lemma}[2][Lemma]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{exercise}[2][Exercise]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{problem}[2][Problem]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{question}[2][Question]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}
\newenvironment{corollary}[2][Corollary]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}\hskip \labelsep {\bfseries #2.}]}{\end{trivlist}}

\newenvironment{solution}{\begin{proof}[Solution]}{\end{proof}}
 
\begin{document}
 
% --------------------------------------------------------------
%                         Start here
% --------------------------------------------------------------
 
\title{Homework 2}
\author{Jia Guo, Georgios Kontoudis\\ 
ME5774 Nonlinear Systems Theory\\
Professor Cornel Sultan} 
\date{Fall 2017}
 
\maketitle
\begin{problem}{1} %theorem, exercise, problem, or question 
Determine if the equilibrium $x_1=x_2=0$ is unstable, stable, asymptotically stable, exponentially stable for
\begin{align*}
&\dot{x_1}=x_2-x_1^3+x_1^5 \\
&\dot{x_2}=-x_1,
\end{align*}
where $\textbf{x}=[x_1 \quad x_2]^\top \in \mathbb{R}^2$. State the strongest claim you can make and provide support for the conclusion. You can consider only local properties.
\end{problem}
\begin{solution}
Let us consider a candidate Lyapunov function
\begin{equation*}
V(x_1,x_2)=\frac{1}{2}x_1^2+\frac{1}{2}x_2^2,
\end{equation*}
that is positive definite $V>0$ for all $x_1,x_2 \in \mathbb{R}^2-\{0 \}$ and $V(0,0)=0$. The rate of change in $V$ yields
\begin{equation*}
\begin{aligned}
& \dot{V}
&  =
&&& x_1\dot{x_1}+x_2\dot{x_2}\\
&& =
&&& x_1(x_2-x_1^3+x_1^5)+x_2(-x_1)\\
&& =
&&& x_1^2(x_1^4-x_1^2).
\end{aligned}
\end{equation*} 
The $x_1,x_2$ have to be confined in $D=\{ \textbf{x} \in \mathbb{R}^2 : -1<x_1<1\}$ that includes the origin, so the function is locally Lipschitz. Then, $x_1^4-x_1^2<0$ for all $x_1,x_2 \in D-\{0 \}$ and $\dot{V}<0$ results a valid Lyapunov function. The origin is an asymptotically stable equilibrium, yet $V$ is radially unbounded,
\begin{equation*}
V(x_1,x_2) \rightarrow \infty \mbox{ as } \| \textbf{x}\| \rightarrow \infty.
\end{equation*}
Thus the origin is a global asymptotically stable equilibrium.
\end{solution}


\begin{problem}{2} %theorem, exercise, problem, or question
Jia Guo
\end{problem}


\begin{problem}{3} %theorem, exercise, problem, or question 
Jia Guo
\end{problem}


\begin{problem}{4} %theorem, exercise, problem, or question 
For
\begin{align*}
&\dot{x_1}=x_1+x_2-x_1(x_1^2+x_2^2) \\
&\dot{x_2}=-x_1+x_2-x_2(x_1^2+x_2^2),
\end{align*}
where $\textbf{x}=[x_1 \quad x_2]^\top \in \mathbb{R}^2$. Show that the equilibrium $x_1=x_2=0$ is unstable using Chetaev's theorem.
\end{problem}
\begin{solution}
For the given system we choose a Lyapunov function candidate
\begin{equation*}
V(x_1,x_2)=\frac{1}{2}x_1^2+\frac{1}{2}x_2^2,
\end{equation*}
which is positive definite $V>0$ for all $x_1,x_2 \in \mathbb{R}-\{0 \}$ and $V(0,0)=0$. The rate of change in $V$ yields
\begin{equation*}
\begin{aligned}
& \dot{V}
&  =
&&& x_1\dot{x_1}+x_2\dot{x_2}\\
&& =
&&& x_1(x_1+x_2-x_1(x_1^2+x_2^2))+x_2(-x_1+x_2-x_2(x_1^2+x_2^2))\\
&& =
&&& (x_1^2+x_2^2)(1-x_1^2-x_2^2).
\end{aligned}
\end{equation*}
Since $x_1^2+x_2^2>0$ for all $\textbf{x}\in  \mathbb{R}^2$, then in order to satisfy $\dot{V}>0$ we need $1-x_1^2-x_2^2>0$ to hold. This means that $\| \textbf{x}\| <1$. Then, we define the set $B_r=\{ \textbf{x}\in \mathbb{R}^2:\| \textbf{x}\| \leq 0.9 \}\subset U$, where $U=\{\textbf{x}\in B_r : V(\textbf{x})>0 \}$. Therefore, by employing Chetaev's theroem we conclode that $\textbf{x}=0 $ is an unstable equilibrium.
\end{solution}

\end{document}


